Which groups are at highest risk for low uptake of flu vaccine?

### 2023-04-23 code/analysis_code/exploratory_analysis.qmd
More updates

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(ggplot2)
library(haven)
library("kableExtra") # For producing a nice table 1
```

@@

### Print data

```{r}
print(mydata2)
```

@@

```{r}
# looking at the data
summary_df <- dplyr::glimpse(mydata2)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```
### Checking names of data
```{r}
names(mydata2)
```

## Data exploration through figures

###Box plots for the flu vaccine uptake and age

```{r}
p1 <- ggplot(mydata2, aes(x=flu_vaccine, y=age)) + 
  geom_boxplot(fill = "gray") +
  labs(title = "Distribution of Age per flu vaccine uptake", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic()

@@

```{r}
figure_file = here("results","age & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

### Barplot for flu vaccine uptake: 1 for no and 2 for yes

```{r}
ggplot(mydata2, aes(flu_vaccine)) + 
  geom_bar(width=0.7, fill="steelblue") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + theme_classic()
```

### Flu vaccine by sex: 1 for no and 2 for yes
```{r}
ggplot(mydata2, aes(x= flu_vaccine, fill = sex)) + 
  geom_bar(position = "dodge") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Male', 'Female')) +  theme_classic()

### Flu vaccine by smoking: 1 for nonsmoker, 2 for former smoker, 3 for current smoker

```{r, message=FALSE, warning=FALSE}
ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Nonsmoker', 'Former Smoker', 'Current smoker')) +
  ggtitle("Distribution of flu vaccine uptake and smoking status") + theme_classic()
```

@@

```{r, message=FALSE, warning=FALSE}
ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Does not have health insurance', 'Does have health insurance')) +
  ggtitle("Distribution of flu vaccine uptake by health insurance status") + theme_classic()
```

@@

## Full table for demogrophics characteristics
```{r, warning=FALSE, message=FALSE}
library(table1)
table_dems <- (table1(~ factor(sex) + age + factor(race) + factor(hispanic) + factor(education) + factor(fam_income) + factor(employ_status) + factor(paid_sick) + factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) + factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status)| flu_vaccine, data=mydata2))

@@

>>>>

```{r, warning=FALSE, message=FALSE}
library(here) #for data loading/saving
library(dplyr)
library(ggplot2)
library(haven)
library(gridExtra) # To combine plots
library("gtsummary") # For producing table 1
#library("kableExtra") # For producing a nice table 1
```

@@

@@

```{r}
# looking at the data
summary_df <- dplyr::glimpse(mydata2)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Checking names of data
```{r}
names(mydata2)
```

# Data exploration through figures

```{r}
getwd()
```

## Box plots for the flu vaccine uptake and age

```{r}
p1 <- ggplot(mydata2, aes(x=flu_vaccine, y=age)) + 
  geom_boxplot(fill = "#A8DADC") + scale_x_discrete(labels = c("No", "Yes")) +
  labs(title = "Figure 1: Distribution of flu vaccine uptake by age", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

@@

### Saving the plot in the result folder
```{r}
figure_file = here("results","Figure 1: Age & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

### Uptake of flu vaccine by race
```{r}
p2 <- ggplot(mydata2, aes(x = flu_vaccine, fill = race)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Figure 2: Distribution of flu vaccine uptake by race",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261", "#E76F51", "#E76F51"), labels = c("White", "Black", "Native American", "Asian")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p2
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","race & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

### Flu vaccine by sex: 1 for no and 2 for yes
```{r}
p3 <- ggplot(mydata2, aes(x = flu_vaccine, fill = sex)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Flu Vaccine Uptake by Sex",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261"), labels = c("Male", "Female")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p3
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","sex & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p3) 
```

### Flu vaccine by smoking: 1 for nonsmoker, 2 for former smoker, 3 for current smoker
```{r}
p4 <- ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Flu Vaccine Uptake by Smoking Status",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261", "#E76F51"), labels = c("Nonsmoker", "Former Smoker", "Current Smoker")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p4
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","smoking status & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p4) 
```

@@

```{r}
p5 <- ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Flu Vaccine Uptake by Insurance Status",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261"), labels = c('Does not have health insurance', 'Does have health insurance')) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p5
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","Insurance status & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p5) 
```

@@

```{r}
tbl_summary(mydata2)
```

```{r}
a <- mydata2 %>% select(age, sex)
b <- tbl_summary(a)
```

## Full table for demogrophics characteristics
```{r, warning=FALSE, message=FALSE}
library(table1)
table_dems <- (table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + chd + alcohol_status + smoking_status| flu_vaccine, data=mydata2))

@@

```{r}
install.packages("kableExtra", dependencies = TRUE)
```

### 2023-04-23 products/manuscript/Manuscript.qmd
More updates

### Description of data and data source

@@

#### Study population

@@

### Statistical analysis

@@

### Exploratory/Descriptive analysis

```{r echo=FALSE, out.width = "50%", fig.align = "center"}
knitr::include_graphics(here("results", "age & flu uptake distribution.png"))
```

@@

### Loading Table 1 from the Exploratory Analysis

@@

### Full analysis

@@

### Model 2: Using all our predictors of interest in the model with flu vaccine uptake as our outcome

@@

### Model Evaluation

#### ROC curve from the full model

@@

#### ROC curve from the reduced model

@@

### Summary and Interpretation

@@

### Strengths and Limitations

@@

### Conclusions

>>>>

## Description of data and data source

@@

## Study population

@@

## Statistical analysis

@@

## Exploratory/Descriptive analysis

```{r echo=FALSE, out.width = "10%",  fig.align = "center"}
knitr::include_graphics(here("results", "age & flu uptake distribution.png"))
```

```{r echo=FALSE, out.width = "10%",  fig.align = "center"}
knitr::include_graphics(here("results", "race & flu uptake distribution.png"))
```

@@

## Loading Table 1 from the Exploratory Analysis

@@

## Full analysis

@@

## Model 2: Using all our predictors of interest in the model with flu vaccine uptake as our outcome

@@

## Model Evaluation

### ROC curve from the full model

@@

### ROC curve from the reduced model

@@

## Summary and Interpretation

@@

## Strengths and Limitations

@@

# Conclusions
